---
title: "Sustained improved survival of patients with metastatic melanoma after the introduction of anti-PD-1-based therapies"
subtitle: <center> Data Load & processing of patient information</center>
author: "Aimilia Schina"
date: '`r paste("First created on May 2022. Updated on ", format(Sys.Date(), "%d %B %Y"))`'
output:
  html_document:
    css: style.css
    code_folding: hide
    #smooth_scroll: yes
    fig_caption: yes
    #highlight: textmate
    #theme: cerulean
    #theme: simplex
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 5
    number_sections: true
    #df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'D:/BIOINF/PROJECTS/2_ANALYSES/18_GIT_PAPERS/RealMelanoma_antiPD1_based')
```


```{r,warning=F, message=F}
RNGkind(sample.kind = "Rounding")
## Clear R-workspace
rm(list=ls(all=TRUE))

## Close all graphic devices
graphics.off()
#####################
### Load packages ###
#####################
library(pacman)
pacman::p_load(extrafont,dplyr, gtsummary, tidyr,data.table, readxl, stringr, tibble, xlsx)
other_pckgs <- c("lubridate")
require(other_pckgs)

extrafont::loadfonts(device="win")
windowsFonts(sans="Palatino Linotype")
loadfonts(device="win")
loadfonts(device="postscript")

```



```{r,warning=F, message=F}
###########################
### Main analysis paths ###
###########################

scriptsPath <- paste0("scripts/")
scriptsFunctionsPath <- paste0(scriptsPath,"functions/")
projectDataPath <- paste0("data/")
projectOutDataPath <- paste0("output/data_files/")
figuresPath <- paste0("output/figures/")
tablesPath <- paste0("output/tables/")
sessionInfoPath <- paste0("session_info/")


##################################
### LOAD SOURCE FUNCTIONS FILE ###
##################################
source(paste0(scriptsFunctionsPath,"data_preprocessing_functions.R"))


#####################
### File suffixes ###
#####################
Rdata.suffix <- ".RData"


```

# Data {.tabset .tabset-fade .tabset-pills}

## Loading patient data

```{r dataLoad,warning=F, message=F,eval=TRUE}
# Loading CSV file with patient info

dataDF.1 <- read_excel(paste0(projectDataPath, "master_12_14_16.xlsx"))
# Remove first empty column
dataDF.1 <- dataDF.1[,-1]
# Select until row 840
dataDF.1 <- dataDF.1[1:839,]
# Correct First column name
colnames(dataDF.1)[1] <- "orig_pt_nt"
colnames(dataDF.1)[2] <- "pt_nt"
colnames(dataDF.1)[c(47,87)] <- c("Comorbidities2", "Comorbidities")
colnames(dataDF.1)[c(55,88)] <- c("Autoimmune2", "Autoimmune")

# Second file for 2018
dataDF.2 <- read_excel(paste0(projectDataPath, "master_18.xlsx"))

# Correct First column name
colnames(dataDF.2)[1] <- "cprnr_hash_md5"
# Add year of diagnosis
dataDF.2 <- dataDF.2 %>% add_column(`Year of diagnosis` = rep(2018,nrow(dataDF.2)), .after=1)
# Here I need correct calculation of age based on date of first visit, since it is rounded for some reason
require(lubridate)
dataDF.2$bday <- str_replace_all(dataDF.2$bday," UTC","")
dataDF.2$`Date-of-first-visit` <- str_replace_all(dataDF.2$`Date-of-first-visit`," UTC","")
dataDF.2$age2 <- get_age(dataDF.2$bday,dataDF.2$`Date-of-first-visit`,dec=TRUE,decPoints = 1)
colnames(dataDF.2)[c(10,172)] <- c("age_rounded","age")
# who--(THIS IS PERFORMANCE STATUS)
# Selected columns for merging both
selColumns <- c("patient_id","Year of diagnosis","gender","age","melanoma_diagnosis","braf","linje1","brainmets","STAGE","ldh_format",
                "who", "Months_to_death_or_last_seen", "Dead_or_alive","PS", "BRAIN", "Steroids", "Comorbidities",
                "Autoimmune", "Other_Malignancies")

# # Check naming of columns to select from both master files
# setdiff(selColumns,colnames(dataDF.2))

# setdiff(selColumns,colnames(dataDF.1))
# [1] "age"                          "braf"                         "brainmets"                    "ldh_format"                  
# [5] "Months_to_death_or_last_seen" "Dead_or_alive"                "Other_Malignancies" 
## Need to change column names here:
# Age column
colnames(dataDF.1)[21] <- "age"
colnames(dataDF.1)[8] <- "braf"
colnames(dataDF.1)[34] <-"brainmets"
colnames(dataDF.1)[27] <- "ldh_format" # also transform to NUMERIC
colnames(dataDF.1)[18] <- "Months_to_death_or_last_seen"
colnames(dataDF.1)[19] <- "Dead_or_alive" # change to NUMERIC
colnames(dataDF.1)[89] <- "Other_Malignancies"

# Transform columns needed in dataDF.1
dataDF.1$ldh_format <- as.numeric(dataDF.1$ldh_format)
dataDF.1$who <- as.numeric(dataDF.1$who)
dataDF.1$Dead_or_alive <- as.numeric(dataDF.1$Dead_or_alive)



dataDF.1.sub <- dataDF.1 %>% dplyr::select(all_of(selColumns)) %>% droplevels()
dataDF.2.sub <- dataDF.2 %>% dplyr::select(all_of(selColumns)) %>% droplevels()


data.DF <- rbind(dataDF.1.sub,dataDF.2.sub)

# Remove rows which are all empty
data.DF <-data.DF[!apply(is.na(data.DF), 1, all), ] 


```

### Fixing first line treatment labels

Looking at first line treatment categories, merging, looking at specific patients

* BMS 047: all patients got a combination of nivolumab + relatlimab
* Nivolumab PD-L1/IDO: all patients got a combination of nivolumab + vaccine anti PD-L1/IDO

We take all patients who got a non-standard treatment, such as the combinations above or 

1. BMS 047
2. Nivo/IDO
3. KN 022 Dabra/Tram + Pembro (which means that this subject actually got a combo of dabrafenib+trametinib+pembrolizumab, this is NOT an approved treatment)  as “experimental”

BMS 511 and BMS 511 Ipi(3)/Nivo(1) vs Ipi(1)/Nivo(3) were assigned to antiPD1+antiCTLA4.


Specific details on assigning treatment:

* Nivolumab PD-L1/IDO --> anti-PD-1: assign to anti-PD-1, even though technically it is experimental 
* BMS 047 Nivolumab/LAG3: Experimental 
* KN 022 Dabra/Tram +/- Pembro:  KN022 to Dabra/Tram. The others are Experimental
* BMS 067 :  It's one patient
* Membrains I:  It's two patients
* Membrains II:  It's one patient: M0172490638
* BMS 047: all patients got a combination of nivolumab + relatlimab
* Nivolumab PD-L1/IDO: all patients got a combination of nivolumab + vaccine anti PD-L1/IDO

All patients who got a non-standard treatment, such as the combinations above or the KN 022 Dabra/Tram + Pembro (which means that this subject actually got a combo of dabrafenib+trametinib+pembrolizumab, this is NOT an approved treatment)  as “experimental”.


```{r ,warning=F, message=F,eval=TRUE}
#issue with anti-PD-1 experimental treats, no treatment

dataDF.new <- data.DF
dataDF.new <- dataDF.new %>% add_column(linje1_v2=dataDF.new$linje1,.after="linje1")

dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("0211237720","0211224920","0111248220","0111245620"))] <-  "Dabrafenib/Trametinib"            
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("M0172514337", "M0172453837"))] <-  "Pembrolizumab"
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("111108120"))] <-  "Nivolumab"
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("0111244220","0211234420","0211261220"))] <-  "KN 022 Dabra/Tram +/- Pembro"                   
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("0211232920"))] <-  ""                      
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("M0172490638"))] <-  "Ipilimumab Nivolumab"
dataDF.new$linje1_v2[which(dataDF.new$patient_id %in% c("111108120"))] <-  "Ipilimumab Nivolumab"# BMS 067


dataDF.new$firstLine <- ifelse(dataDF.new$linje1_v2 %in% c("Pembrolizumab","KN 252 Pembro +/- Epacadostat", "Nivolumab PD-L1/IDO","Nivolumab"), "Anti-PD-1",
            ifelse(dataDF.new$linje1_v2 %in% c("IFN/IL2"), "IFN/IL2",
            ifelse(dataDF.new$linje1_v2 %in% c("Ipilimumab","Ipilimumab (3vs10)","Ipilimumab-IDO"), "Anti-CTLA-4",
            ifelse(dataDF.new$linje1_v2 %in% c("Dabrafenib/Trametinib","Vemurafenib/Cobimetinib", "BRAF/MEK inhib","COMBI-V"), "BRAFi",
            ifelse(dataDF.new$linje1_v2 %in% c("Ipilimumab Nivolumab","BMS 511","BMS 511 Ipi(3)/Nivo(1) vs Ipi(1)/Nivo(3)"), "Anti-PD-1 plus Anti-CTLA-4",
            ifelse(dataDF.new$linje1_v2 %in% c("Temozolomid"), "Temozolomid",
            ifelse(dataDF.new$linje1_v2 %in% c("IL-2 intralesional"), "IL2 intralesional",
            ifelse(dataDF.new$linje1_v2 %in% c("","NULL", NA), "No treatment","Experimental Treatments"))))))))



table(dataDF.new$`Year of diagnosis`)
table(dataDF.new$`Year of diagnosis`, dataDF.new$firstLine)
table(dataDF.new$firstLine)
dim(dataDF.new)
```

### Fixing missing ages

Fix missing ages for 4 patients, based on new information, providing the date of birth of those patients, and the calculation happens as : the 4/5 from 2012, we calculate their age at 1/7-2012 and same for 2016 (at 1/7-2016).

For the following patients, calculate age at first treatment, based on their date of birth as given bellow

* 211235620: 080848 (1st visit: 11/09/2012)
* 111295720: 180356 (1st visit: 24/03/2012)
* 211210620: 240337 (1st visit: 08/10/2012)
* 111253620: 110143 (1st visit: 16/08/2016)

```{r ,warning=F, message=F,eval=TRUE}
# There are values that are zero in age, turn that to missing
dataDF.new$age <- ifelse(dataDF.new$age==0,NA,dataDF.new$age)
dataDF.new[which(is.na(dataDF.new$age)),]

x_date_2012   <- as.Date("2012-09-11")
dataDF.new[which(dataDF.new$patient_id=="0211235620"),]
birth_date <- as.Date("1948-08-08")
dataDF.new$age[which(dataDF.new$patient_id=="0211235620")]<- get_age(birth_date,x_date_2012,dec=TRUE,decPoints = 1)

# 111295720: 180356
x_date_2012   <- as.Date("2012-03-24")
dataDF.new[which(dataDF.new$patient_id=="0111295720"),]
birth_date <- as.Date("1956-03-18")
dataDF.new$age[which(dataDF.new$patient_id=="0111295720")]<- get_age(birth_date,x_date_2012,dec=TRUE,decPoints = 1)

# 211210620: 240337
x_date_2012   <- as.Date("2012-10-08")
dataDF.new[which(dataDF.new$patient_id=="0211210620"),]
birth_date <- as.Date("1937-03-24")
dataDF.new$age[which(dataDF.new$patient_id=="0211210620")]<- get_age(birth_date,x_date_2012,dec=TRUE,decPoints = 1)

# 111253620: 110143
x_date_2016   <- as.Date("2016-08-16")
dataDF.new[which(dataDF.new$patient_id=="0111253620"),]
birth_date <- as.Date("1943-01-11")
dataDF.new$age[which(dataDF.new$patient_id=="0111253620")]<- get_age(birth_date,x_date_2016,dec=TRUE,decPoints = 1)

```


```{r,warning=F, message=F,eval=FALSE}

datatable(dataDF.new, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'excel', 'csv' ),
    scrollX=TRUE,
    pageLength=15
  ),
  caption = 'Patients with other treatments'
)
```

```{r,warning=F, message=F,eval=FALSE}

datatable(dataDF.new %>% dplyr::filter(firstLine=="Experimental Treatments") %>% droplevels(), extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'excel', 'csv' ),
    scrollX=TRUE,
    pageLength=15
  ),
  caption = 'Patients with other treatments'
)
```

### Remove patient

Remove patient 0211237520 that was previously classified as 2016 but was from 2021. It has been removed from master.

## Save merged document for 2012,2014,2016,2018

```{r ,warning=F, message=F,eval=TRUE}

write.xlsx2(dataDF.new %>% as.data.frame(), paste0(projectDataPath, "MERGED_27022023.xlsx"), sheetName = "Sheet1", 
  col.names = TRUE, row.names = FALSE, append = FALSE)
```


# Session info

```{r,warning=F, message=F,eval=TRUE}
session_info <- sessionInfo()
writeLines(capture.output(session_info), paste0(sessionInfoPath,"patient_data_preprocessing.txt"))

sessionInfo()

```
